Interval-Valued Quantum Probabilities: Contextuality and the Born Rule

### 2017-05-12 prop_letter.tex
prop_letter.tex Fix typos proposed in Andy's email...

\documentclass[english,reprint, aps, prl,superscriptaddress, showpacs,
showkeys, longbibliography, amsmath, amssymb]{revtex4-1} 
\usepackage{fullpage}
\usepackage{cmap}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{hyperref}

@@

\begin{abstract}
\begin{description}
\item [{Background}] In a world of bounded resource, any quantum measurement
can be only performed in finite precision.
\item [{Method}] We model the idea of the finite precision measurement
as finite number of intervals, and replace quantum real-valued probability
measures by interval-valued ones.
\item [{Results}] Quantum interval-valued probability measures provide
a unified framework to interpret Gleason's theorem and the Kochen-Specker
theorem as the limiting cases. For the cases between them, we found
some legitimate quantum interval-valued probability measures which
cannot be explained by the Born rule. While restricting on only commuting
projectors, the interval-valued probability measures become classical,
and these non-Born interval-valued probability measures disappear.
\item [{Conclusions}] Our research provides a new insight of finite precision
measurement and non-commutativity when formulating quantum theory.
Also, the Born rule may not be as nature as we used to believe in
the world of finite resource.
\end{description}
\end{abstract}

\pacs{ % https://publishing.aip.org/publishing/pacs/pacs-2010-regular-edition
03.65.Ta, % Foundations of quantum mechanics; measurement theory 
03.67.-a, % Quantum information
02.50.Cw, % Probability theory
02.50.Le} % Decision theory and game theory

@@

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality \cite{Peres2003,BarrettKent2004}.  For example,
Meyer \cite{PhysRevLett.83.3751} and Mermin \cite{Mermin1999} disagree
on whether finite-precision measurement invalidate the Kochen-Specker
theorem \cite{kochenspecker1967,peres1995quantum}, and discussion
continues on whether contextuality is experimentally
verifiable \cite{Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}. The
fundamental question is how to make mathematical theories of quantum
mechanics that are intrinsically, rather than only implicitly,
consistent with the resources available for the realistic accuracy of
an actual measurement.

Our first observation is that essential objectives of our program to
achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number
systems \cite{usat,geometry2013,DQT2014}, may be achievable by
extending classical interval-valued probability
measures \cite{JamisonLodwick2004} to the quantum domain.  We begin by
axiomatizing a quantum interval-valued probability measure (QIVPM)
framework exploiting the fact that a density matrix approach can be
considered as equivalent to a description of quantum states using
probability
measures \cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307
One result is that the expectation value of an observable with respect
to a QIVPM can be defined to be consistent with the constraints on the
expectation value of the same observable expected from continuous
quantum mechanics.  We then examine Gleason's
theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
Kochen-Specker theorem in the context of QIVPM's, and conclude
that they correspond to the limit of an infinite number of precise
intervals, and the set of crypto-deterministic intervals (the set
$\{\imposs,\necess\}$), respectively.

@@

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections of
uncertain events \cite{Kolmogorov1950}.  In the quantum case, the
events of interests are the \emph{projection operators} $P$. This set
includes the empty projection $\mathbb{0}$, the identity
projection $\mathbb{1}$, projections of the form $\proj{\psi}$ where
$\ket{\psi}$ is a pure quantum state, sums of \emph{orthogonal}
projections, and products of \emph{commuting} projections. In a
well-defined quantum probability
space \cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010},
each event $P_i$ is associated with a probability $\mu(P_i)$ subject
to the following axioms: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for a set of
mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$, we
have:
\begin{equation}
\mu\left(\sum_{i=1}^{N}P_{i}\right)=
\sum_{i=1}^{N}\mu\left(P_{i}\right)
\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}

As an example, consider a three-dimensional Hilbert space with
orthonormal basis $\{\ket{0},\ket{1},\ket{2}\}$ and an observable $Q$
with spectral decomposition $\proj{0} + 2\,\proj{1} +
3\,\proj{2}$. Here are two fragments of valid probability measures
$\mu_1$ and $\mu_2$ that can associated with this space:
$\mu_1(\proj{0}) = 0.5$, $\mu_1(\proj{1}) = 0.5$,
$\mu_1(\proj{2}) = 0$, and $\mu_2(\proj{0}) = 0.5$,
$\mu_2(\proj{1}) = 0$, $\mu_2(\proj{2}) = 0.5$. By the Born rule,
the first probability measure corresponds to the quantum system being
in the state $(\ket{0}\pm\ket{1})/\sqrt{2}$ and the second corresponds
to the quantum system being in the state
$(\ket{0}\pm\ket{2})/\sqrt{2}$. The expectation value of the observable
is the first case is 1.5 and in the second it is 2.

@@

To adapt the axioms of such IVPMs to the quantum case, we proceed as
follows. First we fix a collection of intervals $\mathscr{I}$ that
represent an abstraction of the available experimental resources. At
one extreme, infinitely precise resources would correspond to the
collection of intervals $\{[x,x] \mid x \in [0,1]\}$. This case should
coincide with the idealized situation above. At another extreme, an
experiment with \emph{no} resources would correspond to the following
three intervals $\imposs=\left[0,0\right]$,
$\necess=\left[1,1\right]$, and $\unknown=[0,1]$. As before, the first
two intervals are associated with the empty projection and the
identity projection but every other projection would be associated
with $\unknown$; its probability can be anything in the range
$[0,1]$. A third special situation would be to consider IVPMs with
only the two intervals $\imposs=\left[0,0\right]$ and
$\necess=\left[1,1\right]$. This is what Peres calls the
\emph{crypto-deterministic} case in which there is no ambiguity about
the status of any event; we know for certain if it will occur or not.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{prop}
\end{document}

>>>>

\documentclass[english,reprint, aps, prl,superscriptaddress, showpacs,
showkeys, longbibliography, amsmath, amssymb]{revtex4-1} 
\usepackage{fullpage}
\usepackage{cmap}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage[backref=false]{hyperref}

@@

\begin{abstract}~
\begin{description}
\item [{Background}] In a world of bounded resource, any quantum measurement
can be only performed in finite precision.
\item [{Method}] We model the idea of the finite precision measurement
as finite number of intervals, and replace quantum real-valued probability
measures by interval-valued ones.
\item [{Results}] Quantum interval-valued probability measures provide
a unified framework to interpret Gleason's theorem and the Kochen-Specker
theorem as the limiting cases. For the cases between them, we found
some legitimate quantum interval-valued probability measures which
cannot be explained by the Born rule. While restricting on only commuting
projectors, the interval-valued probability measures become classical,
and these non-Born interval-valued probability measures disappear.
\item [{Conclusions}] Our research provides a new insight of finite precision
measurement and non-commutativity when formulating quantum theory.
Also, the Born rule may not be as nature as we used to believe in
the world of finite resource.
\end{description}
\end{abstract}

\pacs{ % https://publishing.aip.org/publishing/pacs/pacs-2010-regular-edition
03.65.Ta, 03.67.-a, 02.50.Cw, 02.50.Le}

@@

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality \cite{Peres2003,BarrettKent2004}.  For example,
Meyer \cite{PhysRevLett.83.3751} and Mermin \cite{Mermin1999} disagree
on whether finite-precision measurements invalidate the Kochen-Specker
theorem \cite{kochenspecker1967,peres1995quantum}, and discussion
continues on whether contextuality is experimentally
verifiable \cite{Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}. The
fundamental question is how to make mathematical theories of quantum
mechanics that are intrinsically, rather than only implicitly,
consistent with the resources available for the realistic accuracy of
an actual measurement.

Our first observation is that essential objectives of our program to
achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number
systems \cite{usat,geometry2013apsrev4,DQT2014}, may be achievable by
extending classical interval-valued probability
measures \cite{JamisonLodwick2004} to the quantum domain.  We begin by
axiomatizing a quantum interval-valued probability measure (QIVPM)
framework exploiting the fact that a density matrix approach can be
considered as equivalent to a description of quantum states using
probability
measures \cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307
One result is that the expectation value of an observable with respect
to a QIVPM can be defined to be consistent with the constraints on the
expectation value of the same observable expected from continuous
quantum mechanics.  We then examine Gleason's
theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
Kochen-Specker theorem in the context of QIVPM's, and conclude
that they correspond to the limit of an infinite number of precise
intervals, and the set of crypto-deterministic intervals (the set
$\{\imposs,\necess\}$), respectively.

@@

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections of
uncertain events \cite{Kolmogorov1950}.  In the quantum case, the
events of interest are the \emph{projection operators} $P$. This set
includes the empty projection $\mathbb{0}$, the identity
projection $\mathbb{1}$, projections of the form $\proj{\psi}$ where
$\ket{\psi}$ is a pure quantum state, sums of \emph{orthogonal}
projections, and products of \emph{commuting} projections. In a
well-defined quantum probability
space \cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010},
each event $P_i$ is associated with a probability $\mu(P_i)$ subject
to the following axioms: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for a set of
mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$, we
have:
\begin{equation}
\mu\left(\sum_{i=1}^{N}P_{i}\right)=
\sum_{i=1}^{N}\mu\left(P_{i}\right)
\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}

As an example, consider a three-dimensional Hilbert space with
orthonormal basis $\{\ket{0},\ket{1},\ket{2}\}$ and an observable $Q$
with spectral decomposition $\proj{0} + 2\,\proj{1} +
3\,\proj{2}$. Here are two fragments of valid probability measures
$\mu_1$ and $\mu_2$ that can associated with this space:
$\mu_1(\proj{0}) = 0.5$, $\mu_1(\proj{1}) = 0.5$,
$\mu_1(\proj{2}) = 0$, and $\mu_2(\proj{0}) = 0.5$,
$\mu_2(\proj{1}) = 0$, $\mu_2(\proj{2}) = 0.5$. By the Born rule,
the first probability measure corresponds to the quantum system being
in the state $(\ket{0}\pm\ket{1})/\sqrt{2}$ and the second corresponds
to the quantum system being in the state
$(\ket{0}\pm\ket{2})/\sqrt{2}$. The expectation value of the observable
in the first case is 1.5 and in the second it is 2.

@@

To adapt the axioms of such IVPMs to the quantum case, we proceed as
follows. First we fix a collection of intervals $\mathscr{I}$ that
represents an abstraction of the available experimental resources. At
one extreme, infinitely precise resources would correspond to the
collection of intervals $\{[x,x] \mid x \in [0,1]\}$. This case should
coincide with the idealized situation above. At another extreme, an
experiment with \emph{no} resources would correspond to the following
three intervals $\imposs=\left[0,0\right]$,
$\necess=\left[1,1\right]$, and $\unknown=[0,1]$. As before, the first
two intervals are associated with the empty projection and the
identity projection but every other projection would be associated
with $\unknown$; its probability can be anything in the range
$[0,1]$. A third special situation would be to consider IVPMs with
only the two intervals $\imposs=\left[0,0\right]$ and
$\necess=\left[1,1\right]$. This is what Peres calls the
\emph{crypto-deterministic} case in which there is no ambiguity about
the status of any event; we know for certain if it will occur or not.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{apsrev4-1}
\bibliography{prop}
\end{document}
